\Large Analyzing the Supreme Court Citation Network

### 2017-10-12 Tex/paper.tex
algorithmics

\documentclass[headsepline=true, abstracton]{scrartcl}
%\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{verbatim}
\usepackage[a4paper,text={140mm,215mm},centering,headsep=10mm,footskip=10mm]{geometry}
\usepackage[urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black,bookmarks]{hyperref}
\usepackage{aliascnt}
\usepackage{lmodern}
\usepackage{mdwlist}
\usepackage{listings}
\usepackage[activate]{pdfcprot}
\usepackage{graphicx}
\usepackage{slashed}
\usepackage{mathrsfs}
\usepackage{geometry}
\usepackage{float}
\usepackage{oldgerm}
\usepackage{setspace}
\usepackage{dsfont}
%\usepackage{mathtools}
\usepackage[all]{xy}
\usepackage{cite}
\usepackage{url}
\pagestyle{headings}
\newcommand{\myclearpage}{\clearpage}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\newenvironment{gelaber}{}{}
\newenvironment{preamble}{}{}
\newcommand{\tostar}{\overset{*}{\lower0.5em\hbox{$\smash{\scriptscriptstyle\rightharpoonup}$}}}
\newtheorem{mydef}{Definition}
\newtheorem{bem}{Bemerkung}
\usepackage{tikz}
\newcommand\circlearound[1]{%
  \tikz[baseline]\node[draw,shape=circle,anchor=base] {#1} ;}

@@

\subsection*{Maximum Likelihood Estimator}
The more rigorous technique is to estimate the parameters directly with the log-likelihood function derived from (\ref{ercm}), which has the following form:
%
\begin{equation}
\text{loglik}(\theta)=\theta^T \cdot h(c(t)| c(t-))-\log(\kappa(\theta))
\label{loglik}
\end{equation}
%
The problem resulting from estimating the parameters with (\ref{loglik}) is that the term
%
$$\kappa(\theta):= \sum_{c(t)^* \in \mathcal{C}(N)} \exp(\theta^T \cdot h(c(t)^*|c(t-)))$$ 
%
which sums up the weighted statistics of all possible binar vectors of length $N$, has to be evaluated. However, the cardinality of $\mathcal{C}(N)$ (\#$(\mathcal{C})=2^N$) is incredibly large and a direkt calculation of this sum is for already small $N$ not feasible. \\[0.3cm]
An solution for this limitation is based on the following consideration: Fix a vector of parameters $\theta_0 \in \Theta$ from the underlying parameter range $\Theta$ and compute for $\theta \in \Theta$ the expected value
%
\begin{eqnarray*}
\mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right]&=& 
\sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot \mathbb{P}_{\theta_0}(C(t)=c(t))\\
&=& \sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot 
\frac{\exp(\theta_0^T \cdot \Gamma(c(t)))}{\kappa(\theta_0)}\\
&=& \frac{1}{\kappa(\theta_0)} \sum_{c(t) \in \mathcal{C}(N)}\exp\left(\theta^T \cdot \Gamma(c(t))\right)\\
&=&\frac{\kappa(\theta)}{\kappa(\theta_0)}
\end{eqnarray*}
%
This equation offers the following possibility: If one draws $L$ random vectors $c^{(1)}(t), \dots ,c^{(L)}(t)$ out of a distribution $\mathbb{P}_{\theta_0}$ appropriately, one gets with the law of big numbers and a big enough sample $L$ the following relation:
%
\begin{equation}
\frac{1}{L}\cdot \sum_{i=1}^{L}  \exp\left((\theta - \theta_0)^T \cdot \Gamma(c^{(i)}(t))\right)
~~\approx~~ \mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right] = \frac{\kappa(\theta)}{\kappa(\theta_0)}
\label{konver}
\end{equation}
%
This approximate can then be used to approximate the log likelihood function.\\[0.4cm]
Next, we will discuss how a sufficient number of suitable drawings $c^{(1)}(t), \dots ,c^{(L)}(t)$ can be sampled from the distribution $\mathbb{P}_{\theta_0}$. \\
For this purpose, the Markov Chain Monte Carlo (MCMC) methods can be used.
\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
To be able to compute the approximate likelihood function one needs a sufficiently large number of random vectors from the distribution $\mathbb{P}_{\theta_0}$. Snijders \cite{Snijders.2002b} introduces an approach to sample random networks for the ERGM framework by using \textit{MCMC methods}. We adapt this approach for sampling appropriate binary vectors for the ERCM. \\[0.4cm]
\textit{Metropolis-Hastings Algorithm}\\\\
Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ to start with (e.g., the observed vector). For $k \in \{0,...,L-1\}$ recursively proceed as follows:\\
\begin{enumerate}
\item Randomly choose a number $i\in \{1,\dots, N\}$
\item Compute, using the equation (\ref{Logit}) the value
$$\pi := \dfrac{\mathbb{P}_{\theta}(C_{i}(t) \neq c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}{\mathbb{P}_{\theta}(C_{i}(t) = c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}$$
\item Define $\delta:= \min\{1, \pi\}$ and draw a random number $Z$ from Bin$(1, \delta)$. If
\begin{itemize}
\item $Z=0$, let $c^{(k+1)}(t) := c^{(k)}(t)$ 
\item $Z=1$, define $c^{(k+1)}(t)$ as
$$c_{p}^{(k+1)}(t)=\begin{cases}
1-c_{p}^{(k)}(t)& \text{if}~ p=i \\
c_{p}^{(k)}(t) &\text{if}~ p \neq i 
\end{cases}$$
\end{itemize}
\item Start at step 1 with $c^{(k+1)}(t)$.
\end{enumerate}
The first $B \ll L$ vectors are discarded as Burn-In.

>>>>

\documentclass[headsepline=true, abstracton]{scrartcl}
%\usepackage[ngerman]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{enumerate}
\usepackage{verbatim}
\usepackage[a4paper,text={140mm,215mm},centering,headsep=10mm,footskip=10mm]{geometry}
\usepackage[urlcolor=black,colorlinks=true,linkcolor=black,citecolor=black,bookmarks]{hyperref}
\usepackage{aliascnt}
\usepackage{lmodern}
\usepackage{mdwlist}
\usepackage{listings}
\usepackage[activate]{pdfcprot}
\usepackage{graphicx}
\usepackage{slashed}
\usepackage{mathrsfs}
\usepackage{geometry}
\usepackage{float}
\usepackage{oldgerm}
\usepackage{setspace}
\usepackage{dsfont}
%\usepackage{mathtools}
\usepackage[all]{xy}
\usepackage{cite}
%\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{algorithm2e}
\usepackage{url}
\pagestyle{headings}
\newcommand{\myclearpage}{\clearpage}
\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
\newenvironment{gelaber}{}{}
\newenvironment{preamble}{}{}
\newcommand{\tostar}{\overset{*}{\lower0.5em\hbox{$\smash{\scriptscriptstyle\rightharpoonup}$}}}
\newtheorem{mydef}{Definition}
\newtheorem{bem}{Bemerkung}
\usepackage{tikz}
\newcommand\circlearound[1]{%
  \tikz[baseline]\node[draw,shape=circle,anchor=base] {#1} ;}

@@

\subsection*{Maximum Likelihood Estimator}
The more rigorous technique is to estimate the parameters directly with the log-likelihood function derived from (\ref{ercm}), which has the following form:
%
\begin{equation}
\text{loglik}(\theta)=\theta^T \cdot h(c(t)| c(t-))-\log(\kappa(\theta))
\label{loglik}
\end{equation}
%
The problem resulting from estimating the parameters with (\ref{loglik}) is that the term
%
$$\kappa(\theta):= \sum_{c(t)^* \in \mathcal{C}(N)} \exp(\theta^T \cdot h(c(t)^*|c(t-)))$$ 
%
which sums up the weighted statistics of all possible binar vectors of length $N$, has to be evaluated. However, the cardinality of $\mathcal{C}(N)$ (\#$(\mathcal{C})=2^N$) is incredibly large and a direkt calculation of this sum is for already small $N$ not feasible. \\[0.3cm]
An solution for this limitation is based on the following consideration: Fix a vector of parameters $\theta_0 \in \Theta$ from the underlying parameter range $\Theta$ and compute for $\theta \in \Theta$ the expected value
%
\begin{eqnarray*}
\mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right]&=& 
\sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot \mathbb{P}_{\theta_0}(C(t)=c(t))\\
&=& \sum_{c(t) \in \mathcal{C}(N)}\exp\left((\theta - \theta_0)^T \cdot \Gamma(c(t))\right)\cdot 
\frac{\exp(\theta_0^T \cdot \Gamma(c(t)))}{\kappa(\theta_0)}\\
&=& \frac{1}{\kappa(\theta_0)} \sum_{c(t) \in \mathcal{C}(N)}\exp\left(\theta^T \cdot \Gamma(c(t))\right)\\
&=&\frac{\kappa(\theta)}{\kappa(\theta_0)}
\end{eqnarray*}
%
This equation offers the following possibility: If one draws $L$ random vectors $c^{(1)}(t), \dots ,c^{(L)}(t)$ out of a distribution $\mathbb{P}_{\theta_0}$ appropriately, one gets with the law of big numbers and a big enough sample $L$ the following relation:
%
\begin{equation}
\frac{1}{L}\cdot \sum_{i=1}^{L}  \exp\left((\theta - \theta_0)^T \cdot \Gamma(c^{(i)}(t))\right)
~~\approx~~ \mathbb{E}_{\theta_0}\left[ \exp\left((\theta - \theta_0)^T \cdot \Gamma(C(t))\right) \right] = \frac{\kappa(\theta)}{\kappa(\theta_0)}
\label{konver}
\end{equation}
%
This approximate can then be used to approximate the log likelihood function.\\[0.4cm]
Next, we will discuss how a sufficient number of suitable drawings $c^{(1)}(t), \dots ,c^{(L)}(t)$ can be sampled from the distribution $\mathbb{P}_{\theta_0}$. \\
For this purpose, the Markov Chain Monte Carlo (MCMC) methods can be used.
\subsection*{Gibbs sampling for the ERCM}
To be able to compute the approximate likelihood function one needs a sufficiently large number of random vectors from the distribution $\mathbb{P}_{\theta_0}$. Snijders \cite{Snijders.2002b} introduces an approach to sample random networks for the ERGM framework by using \textit{MCMC methods}. We adapt this approach for sampling appropriate binary vectors for the ERCM. \\[0.4cm]
\textit{Gibbs sampling}\\
%Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g. observed vector). Afterwards, the length $L$ of the respective sub-sequence is determined. For $k \in \{0,...,L-1\}$ execute the following steps recursively (here the vector in its $k$th iteration is denoted as $c^{(k)}(t)$):\\

\begin{algorithm}[H]
 Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ (e.g. observed vector)\\
 \For{i in 1:N}{
  Compute $\pi:= \cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$\\
  Draw a random number Z from Bin(1,$\pi$)\\
  \eIf{Z=1}{
   set $c^{(k+1)}_i = 1$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
   }{
   set $c^{(k+1)}_i = 0$ and $c^{(k+1)}_j=c^{(k)}_j$, if $i\neq j$
  }
 }
 Start all over using $c^{(k+1)}$\\[0.3cm]
 \caption{Simulation of vectors of $\mathbb{P}_\theta$ using Gibbs sampling}
\end{algorithm}
\vspace{0.5cm}
%\begin{enumerate}
%\item Randomly choose a number $i \in \{1,\dots, N\}$
%\item Compute using the likelihood the value 
%$$\pi:= \mathbb{P}_{\theta}(C_{i}(t)=1 | C_{i}^c(t)=(c_{i}^{(k)}(t))^c)=\cfrac{\exp(\theta^T\cdot \Delta(c_i)(t))}{1+\exp(\theta^T\cdot \Delta(c_i)(t))}$$
%\item Draw a random number $Z$ from Bin$(1, \pi)$. %If
%\begin{itemize}
%\item $Z=0$, define $c^{(k+1)}(t)$ via
%$$c_{p}^{(k+1)}(t)=\begin{cases}
%0& \text{if}~ p=i \\
%c_{p}^{(k)}(t) &\text{if}~ p \neq i 
%\end{cases}$$
%\item $Z=1$, define $c^{(k+1)}(t)$ via
%$$c_p^{(k+1)}(t)=\begin{cases}
%1& \text{if}~ p=i \\
%c_p^{(k)}(t) &\text{if}~ p \neq i 
%\end{cases}$$
%\end{itemize}
%\item Start at step 1 with $c^{(k+1)}(t)$.
%\end{enumerate}
\noindent Using the depicted algorithm, a sequence of random vectors $c^{(0)}(t),...,c^{(L)}(t)$ can be simulated. Since the orignial vector was chosen randomly and the first simulated vectors are very dependent on the chosen mvector (only one entry is changed per iteration!), usually the first $B$ vectors, where $N \ll B \ll L$, are discarded as the so called \textit{Burn-In}.
\subsection*{Metropolis Hastings for the ERCM}\label{networksimulation}
Choose any vector $c^{(0)}(t) \in \mathcal{C}(N)$ to start with (e.g., the observed vector). For $k \in \{0,...,L-1\}$ recursively proceed as follows:\\
\begin{enumerate}
\item Randomly choose a number $i\in \{1,\dots, N\}$
\item Compute, using the equation (\ref{Logit}) the value
$$\pi := \dfrac{\mathbb{P}_{\theta}(C_{i}(t) \neq c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}{\mathbb{P}_{\theta}(C_{i}(t) = c_{i}^{(k)}(t) ~| ~C_i(t)^c=c_i(t)^c)}$$
\item Define $\delta:= \min\{1, \pi\}$ and draw a random number $Z$ from Bin$(1, \delta)$. If
\begin{itemize}
\item $Z=0$, let $c^{(k+1)}(t) := c^{(k)}(t)$ 
\item $Z=1$, define $c^{(k+1)}(t)$ as
$$c_{p}^{(k+1)}(t)=\begin{cases}
1-c_{p}^{(k)}(t)& \text{if}~ p=i \\
c_{p}^{(k)}(t) &\text{if}~ p \neq i 
\end{cases}$$
\end{itemize}
\item Start at step 1 with $c^{(k+1)}(t)$.
\end{enumerate}
The first $B \ll L$ vectors are discarded as Burn-In.
